Incorporating Gyroscope Information in Gesture Recognition

### 2011-12-17 paper/paper.tex
finished bib

\section{Future Research}
We see a variety of promising avenues for the future development of Cluster-Matching. More work needs to be done to assess the scalability of our method, although the outlook is promising. Current literature on accelerometer-based gesture recognition usually reports classification results for 4 to 8 gesture types. We believe that our method has the potential to scale to more gesture types given our current results. Our technique currently incorporates centroid-distance and HMM log-likelihood scores heuristically by simply multiplying them before classifying. It is likely that a more principled method of combining centroid-distance and HMM loglikelihood - one possibly incorporating the data itself - will yield better results. This becomes especially important for gestures in which one component of the classifier performs significantly better than the other. Lastly, future work should address the possibility of using Cluster-Matching for automated segmentation of continuous accelerometer and gyroscope data. Current proposals use only heuristic methods and do not incorporate the trained gesture classifier, e.g. Prekopcsak (2008) demonstrates the feasibility of using a speed threshold to detect the onset of a gesture. HMMs do not provide an obvious means by which to use them for segmentation, as they will readily assign log-likelihoods to sequences of any length. We suspect that Cluster-Matching can be computed over a sliding window on the data stream, providing a means by which to determine when a sequence of samples is being fit well by a subset of the clusters of one of the gesture cluster-sets.

\section{Conclusions}\label{conclusions}
We thus show non trivial gains in accuracy with resect to our baseline classifier KHMM.  These gains are accrued not only through the collection of a new set of features using gyroscope data, but also through our novel algorithms Cluster-Matching-HMM and Cluster-Matching.  Interestingly enough it appears that even though we are classifying sequential data, simply assigning cluster fitness scores to all the vectors in a gesture example, regardless of sequential order, returns the most accurate classifiers.  In fact, the only times when this is not the case is when we synthetically manipulate the data files to purposely break our Cluster-Matching classifier.  However, we could not mimic these this ``break case" using actually gesture data.  We are thus confident in the ability of Cluster-Matching to accurately classify real human gestures.  Regardless, we hope to have pushed forward the study and field of gesture recognition; not only by presenting classifiers with high accuracy, but also through the presentation of a new cluster based paradigm of classifiers.

\bibliographystyle{abbrv}
\bibliography{main}

>>>>

\section{Future Research}
We see a variety of promising avenues for the future development of Cluster-Matching. More work needs to be done to assess the scalability of our method, although the outlook is promising. Current literature on accelerometer-based gesture recognition usually reports classification results for 4 to 8 gesture types. We believe that our method has the potential to scale to more gesture types given our current results. Our technique currently incorporates centroid-distance and HMM log-likelihood scores heuristically by simply multiplying them before classifying. It is likely that a more principled method of combining centroid-distance and HMM loglikelihood - one possibly incorporating the data itself - will yield better results. This becomes especially important for gestures in which one component of the classifier performs significantly better than the other. Lastly, future work should address the possibility of using Cluster-Matching for automated segmentation of continuous accelerometer and gyroscope data. Current proposals use only heuristic methods and do not incorporate the trained gesture classifier, e.g. Prekopcsák demonstrates the feasibility of using a speed threshold to detect the onset of a gesture. HMMs do not provide an obvious means by which to use them for segmentation, as they will readily assign log-likelihoods to sequences of any length [3]. We suspect that Cluster-Matching can be computed over a sliding window on the data stream, providing a means by which to determine when a sequence of samples is being fit well by a subset of the clusters of one of the gesture cluster-sets.

\section{Conclusions}\label{conclusions}
We thus show non trivial gains in accuracy with resect to our baseline classifier KHMM.  These gains are accrued not only through the collection of a new set of features using gyroscope data, but also through our novel algorithms Cluster-Matching-HMM and Cluster-Matching.  Interestingly enough it appears that even though we are classifying sequential data, simply assigning cluster fitness scores to all the vectors in a gesture example, regardless of sequential order, returns the most accurate classifiers.  In fact, the only times when this is not the case is when we synthetically manipulate the data files to purposely break our Cluster-Matching classifier.  However, we could not mimic this ``break case" using actual gesture data.  We are thus confident in the ability of Cluster-Matching to accurately classify real human gestures.  Regardless, we hope to have pushed forward the study and field of gesture recognition; not only by presenting classifiers with high accuracy, but also through the presentation of a new cluster based paradigm of classifiers.

  \begin{thebibliography}{1}

  \bibitem{orig} Klingmann, Marco. {\em Accelerometer-Based Gesture Recognition with the iPhone}. Masters Thesis, Goldsmith University of London. 2009.

  \bibitem{svm} Wu, Jiahui and Pan, Gang and Zhang, Daqing and Qi, Guande and Li, Shijian. {\em Gesture Recognition with a 3-D Accelerometer}.  2009.

  \bibitem{seg}  Zoltán Prekopcsák {\em Accelerometer Based Real-Time Gesture Recognition}. 2008.

  \end{thebibliography}
